% conclusions.tex — Section V: Conclusions
% Target: PRD two-column (revtex4-2)
% Macros used: \HH, \dL, \Mz, \Mbh, \Mstellar, \pdet, \pending

We have investigated whether the redshifted massive black hole mass
$\Mz$, precisely measured from the EMRI gravitational waveform, can
improve dark siren constraints on the Hubble constant $\HH$ with LISA.
The answer is affirmative.  Including $\Mz$ in the Bayesian inference
tightens the $\HH$ posterior by a factor of $\sim 2$, from
$\sim 4\%$ to $\sim 1.8\%$ precision in the $\pdet = 1$ baseline
(\pending{update with production run precision}).  The mechanism is
straightforward: the GW measurement determines $\Mz$ to fractional
precision $\lesssim 10^{-6}$, while the galaxy-side BH mass estimate
from the $\Mbh$--$\Mstellar$ scaling relation carries $\sim 20\%$
uncertainty.  This mismatch in precision acts as a narrow filter on
candidate host galaxies, breaking the $\dL$--$z$ degeneracy that
limits standard dark siren analyses.

As a methodological contribution, we have adapted the catalog
completeness correction of
Gray~\emph{et~al.}~\cite{Gray:2019ksv} --- originally developed for
ground-based dark siren analyses of binary black hole mergers --- to
the EMRI context with LISA and the GLADE+ galaxy
catalog~\cite{Dalya:2021ewn}.  This correction mitigates the
systematic low-$\HH$ bias caused by the preferential cataloging of
low-redshift galaxies (Sec.~\ref{sec:pdet_completeness}).

Several directions for future work will strengthen these results.
A production-scale simulation incorporating the full detection
probability $\pdet(\dL, \Mbh, h)$ and the completeness-corrected
likelihood is underway and will establish the definitive $\HH$
constraint.  Replacing the uniform EMRI population model with an
astrophysically motivated rate model~\cite{Babak:2017tow} will
provide more realistic detection yields.  Finally, marginalizing over
the parameters of the $\Mbh$--$\Mstellar$ scaling
relation~\cite{Reines:2015moa} and extending the analysis to a joint
$(\HH, \Omega_m)$ or $w$CDM inference are straightforward with the
existing pipeline infrastructure.
